---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "John McAvaddy"
output: 
  html_document:
    theme: lumen
    highlight: espresso
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)
```

<br><br>

## Introduction

I chose [data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-09-22) recorded on mountain ascents in the [Himalayan mountain range](https://en.wikipedia.org/wiki/Himalayas). I chose this data because if I could get paid to climb mountains, I would climb one every day for the rest of my life. Because of my love for climbing mountains, I thought it would be interesting to explore climbing data collected from some of the [tallest mountains](https://en.wikipedia.org/wiki/List_of_highest_mountains_on_Earth) in the world. 

```{r read-examine-data}
members <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/members.csv')
expeditions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/expeditions.csv')
peaks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/peaks.csv')

glimpse(members) # see data in members data set
glimpse(expeditions) # see data in expeditions data set
glimpse(peaks) # see data in peaks data set
```

<br><br>

## Question 1

#### Is there a relationship between the average age of an expedition group and whether they successfully reached the peak?

### Methods
```{r question_one_methods}
# members %>%
#   group_by(expedition_id) %>% # groups rows by similar values in 
#   summarize(average_age = mean(age)) %>% # create new column of average ages of each expedition group
#   drop_na() %>% # drop NA here so I don't consider any groups that do not have a "complete" average age
#   left_join(members) %>% # combine average_age column into original data set 
#   mutate(success = if_else(success == TRUE, 1, 0)) %>%
#   select(expedition_id, average_age, success) %>%
#   print()

# members %>%
#   select(expedition_id) %>%
#   distinct() %>%
#   print()

expeditions %>%
  mutate(termination_reason = str_replace(termination_reason, "Success (main peak) ", "Success")) %>%
  mutate(termination_reason = str_replace(termination_reason, "Success (subpeak) ", "Success")) %>%
  select(termination_reason) %>%
  print(n = 100)

members %>%
  filter(expedition_id == "AMAD61101") 


```
### Answer

Replace this text with your answer to your question.


<br><br>

## Question 2

#### Are injuries more likely to occur in the upper half or lower half of the peak's height?
### Methods

### Answer

Replace this text with your answer to your question.

<br><br>

## Question 3

#### In which season are climbers most likely to succeed?

### Methods

### Answer

Replace this text with your answer to your question.


<br><br>

## Question 4

#### What is the minimum average age of a first ascent expedition group?

### Methods

### Answer

Replace this text with your answer to your question.
